---
title: "Estimating Exam Scores: Supervised Learning"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
author: Timmy Miller, MPS Candidate
output:
    html_document:
        css: exams.css
        toc: true
        toc_float:
            collapsed: false
            smooth_scroll: false
        toc_depth: 2
        number_sections: false
---
```{r, include = FALSE}
library(tidyverse)
library(broom)
library(ggfortify)

exams <- read_csv("exams.csv")
```


---
# Introduction
Hi there! My name is Timmy and in this document I will be giving a walk-through on how to run basic
supervised machine learning techniques on a dataset. The purpose of this document is to act as a guide to
someone looking to run a supervised machine learning prediction on multiple variables.

## Dataset
[Link to Data](https://www.kaggle.com/datasets/whenamancodes/students-performance-in-exams)
This dataset tracks the grades for 1000 students on 3 different tests: math, reading, and writing.
The dataset also includes other categorical variables, including things like ethnicity and gender.  

## Creating Dummy Variables
Using the `glimpse()` code from `dplyr`, we can obtain the structure of the exam dataset.
```{r, echo = FALSE}
exams <- exams %>%
    rename(
        "race_ethnicity" = "race/ethnicity",
        "parent_edu" = "parental level of education",
        "math_score" = "math score",
        "read_score" = "reading score",
        "write_score" = "writing score",
        "test_prep_course" = "test preparation course"
    )
glimpse(exams)
```

We can see that we have three numerical variables, `math_score`, `read_score`, and `write_score`.
However, the rest of the variables are character or categorical values. It will be important to transform this categorical
variables into "Dummy Variables", which means replacing each category with a numerical value. 
The code I used to create these dummy variables for each category is shown below.

* **Gender**
    + Male = 1, Female = 0
* **Race/Ethnicity**
    + groupA = 1, groupB = 2, groupC = 3, groupD = 4, groupE = 5
* **Parent Level of Education**
    + some high school = 1, high school = 2, some college = 3, associate's degree = 4, bachelor's degree = 5, master's degree = 6
* **Lunch**
    + standard = 1, reduced/free = 0


### Gender
```{r Gender}
exams_clean <- exams
exams_clean$gender <- ifelse(exams_clean$gender == "male", 1, 0) 
```
### Race/Ethnicity
```{r Race-Ethnicity}
exams_clean$race_ethnicity <- ifelse(exams_clean$race_ethnicity == "group A", 1, exams_clean$race_ethnicity)
exams_clean$race_ethnicity <- ifelse(exams_clean$race_ethnicity == "group B", 2, exams_clean$race_ethnicity)
exams_clean$race_ethnicity <- ifelse(exams_clean$race_ethnicity == "group C", 3, exams_clean$race_ethnicity)
exams_clean$race_ethnicity <- ifelse(exams_clean$race_ethnicity == "group D", 4, exams_clean$race_ethnicity)
exams_clean$race_ethnicity <- ifelse(exams_clean$race_ethnicity == "group E", 5, exams_clean$race_ethnicity)

exams_clean$race_ethnicity <- as.numeric(exams_clean$race_ethnicity)
```
### Parental Level of Education
```{r Parent-Level-Education}
exams_clean$parent_edu <- ifelse(exams_clean$parent_edu == "some high school", 1, exams_clean$parent_edu)
exams_clean$parent_edu <- ifelse(exams_clean$parent_edu == "high school", 2, exams_clean$parent_edu)
exams_clean$parent_edu <- ifelse(exams_clean$parent_edu == "some college", 3, exams_clean$parent_edu)
exams_clean$parent_edu <- ifelse(exams_clean$parent_edu == "associate\'s degree", 4, exams_clean$parent_edu)
exams_clean$parent_edu <- ifelse(exams_clean$parent_edu == "bachelor\'s degree", 5, exams_clean$parent_edu)
exams_clean$parent_edu <- ifelse(exams_clean$parent_edu == "master\'s degree", 6, exams_clean$parent_edu)

exams_clean$parent_edu <- as.numeric(exams_clean$parent_edu)
```
### Lunch
```{r}
exams_clean$lunch <- ifelse(exams_clean$lunch == "standard", 1, 0)
exams_clean$lunch <- as.numeric(exams_clean$lunch)
```
### Test Prep Course 
```{r}
exams_clean$test_prep_course <- ifelse(exams_clean$test_prep_course == "completed", 1, 0)
exams_clean$test_prep_course <- as.numeric(exams_clean$test_prep_course)
```
## Linear Model: Cumulative Exam Score Predictions
We can quickly look at the `structure()` of our `exams_clean` dataset and see our new numerical dummy variables.
```{r structure-new}
str(exams_clean)
```

In this first step, we would like to see if we can predict what a person would obtain as a cumulative score over all three exams. 
In other terms, `cum_score = math_score + read_score + write_score`. Before we can create a liner model, we must create a new 
In R, we will be able to achieve this via the code:
```{r cum-score}
df1 <- exams_clean %>%
    mutate(
        cum_score = math_score + read_score + write_score
    )
df1 <- df1[-c(6:8)] # Removes individual scores
```

Now we would like to create a train set (set that we will train our model on), and a test set (will test the accuracy of our model). We can create the test set by calling a subset of the main dataset.
In our case, we will just take out 100 'random' observations. Here is the code that I used in order to do so:

```{r train-test1}
test <- df1[250:350,]
train <- df1[-c(250:350),]
```

Now, we can call the `lm()` function to create a linear regression with cumulative score as the response variable and all other variables as the independent variables.
In the same chunk, we will also run the `summary()` function to get a good look at our new model.
```{r lm1}
lm1 <- lm(cum_score ~ ., data = train)
summary(lm1)
```

One more important step is to ensure that their are no datapoints that are negatively affecting the 
accuracy of our model due to high leverage or residuals. We can use the `autoplot()`
function of the `ggfortify` package to create a 4x4 grid of plots to help us visualize our model.

```{r autoplot, out.width = '50%', out.width= '.5', fig.cap = "Autoplot of lm1"}
autoplot(lm1)
```
Our plots show that there are no high leverage or residual points.
## Linear Model: Individual Exam Score Predictions

### Math
### Reading
### Writing

## Training a Cumulative Score Model using the caret package

